A Standard for the Scholarly Citation of Archaeological Data

### 2017-09-13 analysis/paper/paper.Rmd
add para on repro and open materials with DOI for this paper

Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)). Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data. In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process. On this last point, Table \@ref(tab:desirable) shows some of the current norms on this, based on Stodden’s (2009a) detailed legal discussion. See also Stodden’s (2009b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs) for a convenient and robust copyright schema for typical research products.

@@

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.) [archaeology data service manuals]. Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

@@

The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats. 'Creator' is equivalent to the author or primary collector and arranger of the data. The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms. The identifier is a special type of persistent internet address. The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes. There are several varieties of these in use, with DOIs (digital object identified) being the most common form. When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.

@@

# Summary

@@

Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S., ... & Errington, T. M. (2016). Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency. PLoS Biol, 14(5), e1002456.

@@

>>>>

Stodden (2010) surveyed registrants at a top machine learning conference (the Neural Information Processing Systems (NIPS) conference) to identify the factors that affect researchers’ decisions to reveal data (Table \@ref(tab:costsbenefits)). Stodden found that communitarian factors influence scientists to share their data, but private incentives determine when scientists choose not to share their code and data. In the top three reasons not to share, we see a lack of incentives to invest time in data sharing (however, there are many other time-consuming tasks in research, and they get done because clear reward structures are in place); a lack of faith in data citation (which we hope this paper will help with); and a lack of an efficient copyright process. Stodden’s (2009a, b) ‘Reproducible Research Standard’, summarised in Figure \@ref(fig:stodden-rrs), is a convenient and robust copyright schema that is in wide use for many typical research projects.

@@

## Reproducibility and open source materials for this study

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in Marwick (2017), we include the entire R code used for all the analysis and visualizations contained in this paper in our SOM at http://dx.doi.org/10.17605/OSF.IO/KSRUZ. Also in this version-controlled compendium are the raw data for all the tests reported here. All of the figures, tables and statistical test results presented here can be independently reproduced with the code and data in this repository. In our SOM our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use (for more details, see Marwick 2017).

@@

In the pilot studies reported above we see three common approaches to sharing data: sending files privately by email, following a personal request; including data as supplementary files submitted for publication with a journal article; and sharing data by depositing files in a trustworthy, public, DOI-issuing repository. The third method, via a public repository, is by far the most widely recommended method both specifically by archaeologists (Kintigh 2006; Kansa et al. 2011; Kansa and Kansa 2013), and in other sciences (cf. Cranston et al. 2014; Penev et al. 2017; Mounce 2014; Thessen & Patterson 2011; Whitlock 2011).	There are many guidelines and conventions from other disciplines that are relevant to sharing archaeological data (ICPSR 2012; Strasser et al. 2012; Wilkinson et al. 2016). Table \@ref(tab:desirable) summarises some of these, and recommendations specifically written for archaeologists can be found in UK Archaeology Data Service (n.d.), ARCHES project. (n.d.). Many of these guidelines may create new technical and practical challenges to archaeologists who are not accustomed to sharing data, and many not want to dedicate resources to sharing data. However, we should not let the perfect be the enemy of the good. In our own work we try to follow these guidelines with no dedicated resources for data sharing, and with an incomplete knowledge of some of the technical details of repository operations. While we recognise that our data sharing pracitses fall short of all the guidelines, we still see value in sharing data imperfectly, rather than not at all.

@@

The 'PublicationYear', 'Title' and 'Publisher' are familiar from traditional citation formats. 'Creator' is equivalent to the author or primary collector and arranger of the data. The 'Version' and 'Identifier' fields are unique to data citation, and can take a variety of forms. The identifier is a special type of persistent internet address. The identifier must have two important qualities: the address contains some metadata (such as the identity of the publisher), and they remained fixed to a specific file or bundle of files over the lifetime of those objects, even if their location on the internet changes. There are several varieties of these in use, with DOIs (digital object identifiers) being the most common form. When a data publisher changes the internet address of a data set, the are responsible for informing the DOI-issuing authority so that the DOI resolves to the new location of the files.

@@

# Conclusion

@@

Kidwell, M. C., Lazarević, L. B., Baranski, E., Hardwicke, T. E., Piechowski, S., Falkenberg, L. S., & Errington, T. M. (2016). Badges to acknowledge open practices: A simple, low-cost, effective method for increasing transparency. PLoS Biol, 14(5), e1002456.

@@

Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation. Journal of Archaeological Method and Theory, 24(2), 424-450.
